---
title: "Reproducible Research - Project 1"
output: html_document
---

## Assignment
###Loading and preprocessing the data. 
1. Load the data (i.e. read.csv())
2. Process/transform the data (if necessary) into a format suitable for your analysis

```{r loadData, echo = TRUE}
library(utils)
library(RCurl)
library(lubridate)
library(dplyr)
temp <- tempfile()
urlfile<-'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'
download.file(urlfile,temp, mode = "wb") # download the file.
data<-read.csv(unzip(temp)) # read unzipped  csv file
str(data)
data$date <- ymd(data$date) # convert date column into date format
```

### What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day
2. If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day
3. Calculate and report the mean and median of the total number of steps taken per day

```{r dataSummary, echo=TRUE}
data2 <- data[is.na(data$steps),] # delete missing data



```